---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
devtools::load_all()
```

# tidyquant

[![Travis-CI Build Status](https://travis-ci.org/mdancho84/tidyquant.svg?branch=master)](https://travis-ci.org/mdancho84/tidyquant)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/tidyquant)](https://cran.r-project.org/package=tidyquant)
<!-- ![](http://cranlogs.r-pkg.org/badges/tidyquant?color=brightgreen)  -->
<!-- ![](http://cranlogs.r-pkg.org/badges/grand-total/tidyquant?color=brightgreen) -->

> Bringing financial analysis to the tidyverse

`tidyquant` integrates the best resources for collecting and analyzing financial data, `zoo`, `xts`, `quantmod` and `TTR`, with the tidy data infrastructure of the `tidyverse` allowing for seamless interaction between each. You can now perform complete financial analyses in the `tidyverse`. 

## Benefits

__The `tidyquant` philosophy:__

* __A few core functions with a lot of power, that__
* __leverage the quantitative analysis power of `zoo`, `xts`, `quantmod` and `TTR`, and are__
* __designed to be used and scaled with the `tidyverse`.__

## Installation

Development Version with Latest Features:

``` {r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("mdancho84/tidyquant")
```

CRAN Approved Version: 

```{r, eval = FALSE}
install.packages("tidyquant")
```


## Examples

Start by loading `tidyquant`.

```{r, eval = FALSE}
# Loads tidyquant, tidyverse, lubridate, quantmod, TTR, and xts/zoo
library(tidyquant) 
```


### Getting Data

`tq_get()` is the one-stop shop for retrieving data. The full list of get options are:

```{r}
tq_get_options()
```

__Stock Prices__:

Set `get = "stock.prices"` to get stock prices. Notice the output is _always_ a `tibble`.

```{r}
aapl_prices <- tq_get("AAPL", get = "stock.prices")
aapl_prices
```

Get data for multiple stocks:

```{r}
tq_get(c("AAPL", "GOOG", "FB"), get = "stock.prices")
```


__Financial Statements__:

Set `get = "financials"` to get financial statements. The statements are returned as nested tibbles, that can be unnested and analyzed together.

```{r}
tq_get("AAPL", get = "financials")
```

Get financials for multiple stocks:

```{r}
tq_get(c("AAPL", "GOOG", "FB"), get = "financials")
```


__Key Ratios__:

Set `get = "key.ratios"` to get 10 years of 89 different key ratios (e.g. P/E, P/S, EPS, ROA, ROE, current ratio, debt/equity, inventory turnover, and many more), separated into seven primary sections. 

```{r}
tq_get("AAPL", get = "key.ratios")
```

Get key ratios for multiple stocks:

```{r}
tq_get(c("AAPL", "GOOG", "FB"), get = "key.ratios")
```


__Stock Indexes__:

There are `r tq_get_stock_index_options() %>% length()` indexes available to select from, which can be seen with a call to `tq_get_stock_index_options()`. 

```{r}
tq_get_stock_index_options()
```

Set `x` to one of the options and `get = "stock.index"` to get a full list of stocks within the specified index. 
```{r}
tq_get("SP500", get = "stock.index")
```

Combine `tq_get` stock index and stock price options to get the prices for __every stock in an index__. _Note that this may take several minutes due to the number of stock prices being generated (results not shown because of this)._

```{r, eval = F}
sp_500_prices <- tq_get("SP500", get = "stock.index") %>%
    tq_get(get = "stock.prices")
```


__Other Options__:

There are many other get options including __dividends__, __splits__, __economic data__ from the FRED, and __exchange rates__ and __metal prices__ from Oanda.

### Working in the tidyverse

You may already know and love `tidyverse` packages like `ggplot2`, `dplyr`, `tidyr`, `purrr`, `readr`, and `tibble` along with `lubridate` for working with date and datetime. `tidyquant` works solely in tibbles, so all of the `tidyverse` functionality is intact. 

A simple example inspired by [Kan Nishida's blog](https://blog.exploratory.io/introducing-time-series-analysis-with-dplyr-60683587cf8a#.w6pvyi3d2) shows the `dplyr` and `lubridate` capability: Say we want the growth in the stock over the past year. We can do this with `dplyr` operations. 

Getting the last year is simple with `dplyr` and `lubridate`. We first `select` the date and adjusted price (adjusted for stock splits). We then `filter` using `lubridate` date functions. We can use the `mutate` to add columns to the data frame: Add the baseline price using the `first` function, add the growth and growth percent versus baseline columns using standard mathematical operations. We tack on a final `select` statement to remove unnecessary columns. The final workflow looks like this:

```{r}
aapl_prices %>%
    select(date, adjusted) %>%
    filter(date >= today() - years(1)) %>%
    mutate(baseline = first(adjusted),
           growth = adjusted - baseline,
           growth_pct = growth / baseline * 100) %>%
    select(-(baseline:growth))
```

### Transforming & Mutating Data with zoo, xts, quantmod, and TTR Functions

You may already know and love `zoo`, `xts`, `quantmod`, and `TTR`, which is why the core functionality is fully integrated. The workhorse functions, `tq_transform()` and `tq_mutate()`, apply `zoo`, `xts`, `quantmod`, and `TTR` functions to `tibbles`. The full list of compatible functions are shown with a call to `tq_transform_fun_options()`. Remove the `%>% str()` to expand the list.

```{r}
tq_transform_fun_options() %>% str()
```


#### tq_transform

`tq_transform()` returns a new data set that can either be in the same periodicity or a different periodicity as the original data set. Let's use `tq_transform` to transform the periodicity of the `aapl_prices`. The `quantmod` OHLC codes are used to select the open, high, low, close, and volume (OHLCV) columns, which are then sent to the transformation function, `xts::to.period`, for transformation to monthly periodicity. We now have a much smaller data set containing the monthly prices. 

```{r}
aapl_prices %>%
    tq_transform(ohlc_fun = OHLCV, transform_fun = to.period, period = "months")
```


#### tq_mutate

The cousin of `tq_transform()` is `tq_mutate()`. While `tq_transform()` produces a new, transformed data set, `tq_mutate()` modifies the existing data set. This is very useful for applying `TTR` functions like `BBands`, `MACD`, Moving Averages, etc. There is one caveat: the mutation must be in the same periodicity as the original data set (otherwise you can't add columns because the rows will not match up). Let's use `tq_mutate()` to add some Bollinger Bands and MACD using the closing prices (Cl in OHLC notation).

```{r}
aapl_prices %>%
    tq_mutate(ohlc_fun = Cl, mutate_fun = MACD) %>%
    tq_mutate(ohlc_fun = HLC, mutate_fun = BBands)
```

#### tq_tranform_xy and tq_mutate_xy 

The "xy" variants are useful in situations where (1) you have two inputs (hence x and y) that don't fit into the OHLC function, or (2) you are working with a single column of non-OHLC data.

__Two inputs that don't fit OHLC mold__:

```{r}
aapl_prices %>%
    tq_mutate_xy(x = close, y = volume, mutate_fun = EVWMA, col_rename = "EVWMA")
```

__Working with a single column of non-OHLC data__:

```{r}
tq_get("GDPC1", get = "economic.data") %>%
    tq_mutate_xy(x = price, mutate_fun = rollapply, width = 5, FUN = mean)
```


### Scaling with the tidyverse

All functions return data sets as `tibbles`, which allows for interaction within the `tidyverse`. This means we can:

* Use `dplyr` and `tidyr` to select, filter, nest/unnest, group_by, etc.
* Use the pipe (`%>%`) for chaining operations.
* Seamlessly scale data retrieval and transformations/mutations using `purrr` to map functions or `dplyr` to combine `group_by` with `tq_mutate` or `tq_transform`.

#### Getting Financial Data for Multiple Stocks

A very basic example is retrieving the stock prices for multiple stocks. There are three primary ways to do this:

__Method 1: Map a character vector with multiple stock symbols__

```{r}
c("AAPL", "GOOG", "FB") %>%
    tq_get(get = "stock.prices", from = "2016-01-01", to = "2017-01-01")
```

The output is a single level tibble with all or the stock prices in one tibble.

__Method 2: Map a tibble with stocks in first column__

First, obtain a tibble of stocks. The stock symbols must be in the first column.

```{r}
stock_list <- tibble(symbols = c("AAPL", "JPM", "CVX"),
                     industry = c("Technology", "Financial", "Energy"))
stock_list
```

Second, send the stock list to `tq_get`. Notice how the symbol and industry columns are expanded the length of the stock prices.


```{r}
stock_list %>%
    tq_get(get = "stock.prices", from = "2016-01-01", to = "2017-01-01")
```


__Method 3: Use purrr to map a function__

We can pipe a tibble of stock symbols to a mutation that maps the `tq_get(get = "stock.prices")` function. The result is all of the stock prices in nested format.

```{r}
tibble(symbol = c("AAPL", "GOOG", "AMZN", "FB", "AVGO", "SWKS","NVDA")) %>%
    mutate(stock.prices = map(.x = symbol, ~ tq_get(.x, get = "stock.prices")))
```

#### Manipulating Financial Data for Multiple Stocks

Once you get the data, you typically want to do something with it. You can easily do this at scale. Let's get the yearly returns for multiple stocks using `tq_transform`. First, get the prices. Second, use `group_by` to group by stock symbol. Third, apply the transformation. We can do this in one easy workflow:

```{r}
c("AAPL", "GOOG", "FB") %>%
    tq_get(get = "stock.prices") %>%
    group_by(symbol.x) %>%
    tq_transform(Ad, transform_fun = periodReturn, period = "yearly")
```


## Further Information

This just scratches the surface of the features. See the [`tidyquant` vignette](https://cran.r-project.org/web/packages/tidyquant/vignettes/tidyquant.html) for further details on the package.

