---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
devtools::load_all()
```

# tidyquant

[![Travis-CI Build Status](https://travis-ci.org/mdancho84/tidyquant.svg?branch=master)](https://travis-ci.org/mdancho84/tidyquant)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/tidyquant)](https://cran.r-project.org/package=tidyquant)
<!-- ![](http://cranlogs.r-pkg.org/badges/tidyquant?color=brightgreen)  -->
<!-- ![](http://cranlogs.r-pkg.org/badges/grand-total/tidyquant?color=brightgreen) -->

`tidyquant` integrates the best quantitative resources for collecting and analyzing quantitative data, `xts`, `quantmod` and `TTR`, with the tidy data infrastructure of the `tidyverse` allowing for seamless interaction between each and working within the `tidyverse`. 

## Benefits

__The `tidyquant` philosophy:__

* __A few core functions with a lot of power, that__
* __leverage the quantitative analysis power of `xts`, `quantmod` and `TTR`, and are__
* __designed to be used and scaled with the `tidyverse`.__

## Installation

To install from CRAN:

```{r, eval = FALSE}
install.packages("tidyquant")
```


Or, to install development version from GitHub:

``` {r}
# install.packages("devtools")
devtools::install_github("mdancho84/tidyquant")
```


## Examples

```{r, eval = FALSE}
# Loads tidyquant, tidyverse, lubridate, quantmod, TTR, and xts
library(tidyquant) 
```


### Getting Data in Tibble Format:

`tq_get()` is the one-stop shop for retrieving data. The full list of get options are:

```{r}
tq_get_options()
```

Set `get = "stock.prices"` to get stock prices. Notice the output is _always_ a `tibble`.

```{r}
aapl_prices <- tq_get("AAPL", get = "stock.prices")
aapl_prices
```

Set `get = "financials"` to get financial statements. The statements are returned as nested tibbles, that can be unnested and analyzed together.

```{r}
tq_get("AAPL", get = "financials")
```

There are many other get options including stock indexes, dividends, splits, economic data from the FRED, and exchange rates and metals from Oanda.

### Working in the tidyverse

You probably already know and love `tidyverse` packages like `dplyr`, `tidyr`, `purrr`, `readr`, and `tibble` along with `lubridate` for working with date and datetime. `tidyquant` works solely in tibbles, so all of the `tidyverse` functionality is intact. 

A simple example inspired by [Kan Nishida's blog](https://blog.exploratory.io/introducing-time-series-analysis-with-dplyr-60683587cf8a#.w6pvyi3d2) shows the `dplyr` and `lubridate` capability: Say we want the growth in the stock over the past year. We can do this with `dplyr` operations. 

Getting the last year is simple with `dplyr` and `lubridate`. We first `select` the date and adjusted price (adjusted for stock splits). We then `filter` using `lubridate` date functions. We can also get a baseline price using the `first` function. Growth and growth percent versus baseline columns can be added now. We tack on a final select statement to remove unnecessary columns. The final workflow looks like this:

```{r}
aapl_prices %>%
    select(date, adjusted) %>%
    filter(date >= today() - years(1)) %>%
    mutate(baseline = first(adjusted),
           growth = adjusted - baseline,
           growth_pct = growth / baseline * 100) %>%
    select(-(baseline:growth))
```

### Transforming & Mutating Data with xts, quantmod, and TTR Functions

You may already know and love `xts`, `quantmod`, and `TTR`, which is why the core functionality is fully integrated. The workhorse functions are `tq_transform()` and `tq_mutate()`. These functions leverage the power of `xts`, `quantmod`, and `TTR`. The full list of `xts`, `quantmod`, and `TTR` functions that can be used are:

```{r}
tq_transform_fun_options()
```


#### tq_transform

`tq_transform()` returns a new data set that can either be in the same periodicity or a different periodicity as the original data set. Let's use `tq_transform` to transform the periodicity of the `aapl_prices`. The `quantmod` OHLC codes are used to select the open, high, low, close, and volume (OHLCV) columns, which are then sent to the transformation function, `xts::to.period`, for transformation to monthly periodicity. We now have a much smaller data set containing the monthly prices. 

```{r}
aapl_prices %>%
    tq_transform(x_fun = OHLCV, transform_fun = to.period, period = "months")
```


#### tq_mutate

The cousin of `tq_transform()` is `tq_mutate()`. While `tq_transform()` produces a new, transformed data set, `tq_mutate()` modifies the existing data set. This is very useful for applying `TTR` functions like `BBands`, `MACD`, Moving Averages, etc. There is one caveat: the mutation must be in the same periodicity as the original data set (otherwise you can't add columns because the rows will not match up). Let's use `tq_mutate()` to add some Bollinger Bands and MACD using the closing prices (Cl in OHLC notation).

```{r}
aapl_prices %>%
    tq_mutate(x_fun = Cl, mutate_fun = MACD) %>%
    tq_mutate(x_fun = HLC, mutate_fun = BBands)
```


### Scaling with the tidyverse

All functions return data sets as `tibbles`, which allows for interaction within the `tidyverse`. This means we can:

* Use `dplyr` and `tidyr` to select, filter, nest/unnest, etc.
* Use the pipe (`%>%`) for chaining operations.
* Seamlessly scale data retrieval and transformations/mutations using `purrr` to map functions.

A very basic example is retrieving the stock prices for multiple stocks. We can do this by piping a tibble of stock symbols to a mutation that maps the `tq_get(get = "stock.prices")` function.

```{r}
tibble(symbol = c("AAPL", "GOOG", "AMZN", "FB", "AVGO", "SWKS","NVDA")) %>%
    mutate(stock.prices = map(.x = symbol, ~ tq_get(.x, get = "stock.prices")))
```


## Further Information

This just scratches the surface of the features. See the [`tidyquant` vignette](vignettes/tidyquant.md) for further details on the package.

