---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
devtools::load_all()
```

# tidyquant

[![Travis-CI Build Status](https://travis-ci.org/mdancho84/tidyquant.svg?branch=master)](https://travis-ci.org/mdancho84/tidyquant)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/tidyquant)](https://cran.r-project.org/package=tidyquant)


`tidyquant` integrates the best quantitative resources for collecting and analyzing quantitative data, `xts`, `quantmod` and `TTR`, with the tidy data infrastructure of the `tidyverse` allowing for seamless interaction between each. 

## Benefits

__The `tidyquant` philosophy:__

* __A few core functions with a lot of power, that__
* __leverage the quantitative analysis power of `xts`, `quantmod` and `TTR`, and are__
* __designed to be scaled with the `tidyverse` workflow.__

## Installation

To install from GitHub:

```R
# install.packages("devtools")
devtools::install_github("mdancho84/tidyquant")
```


## Examples

```{r, eval = FALSE}
library(tidyquant) # Loads tidyquant, tidyverse, quantmod, TTR, and xts
```


### Getting Data (tq_get):

`tq_get()` is the one-stop shop for retrieving data. The full list of get options are:

```{r}
tq_get_options()
```

Set `get = "stock.prices"` to get stock prices. Notice the output is always a `tibble`.

```{r}
aapl_prices <- tq_get("AAPL", get = "stock.prices")
aapl_prices
```

Set `get = "financials"` to get financial statements.

```{r}
tq_get("AAPL", get = "financials")
```




### Transforming & Mutating Data (tq_transform and tq_mutate):

The workhorse functions are `tq_transform()` and `tq_mutate()`, which leverage the power of `xts`, `quantmod`, and `TTR`. The full list of functions that can be used are:

```{r}
tq_transform_fun_options()
```


#### tq_transform

First, let's use `tq_transform` to transform the periodicity of the `aapl_prices`. The `quantmod` OHLC codes are used to select the open, high, low, close, and volume (OHLCV) columns, which are then sent to the transformation function, `xts::to.period`, for transformation to monthly periodicity. We now have a much smaller data set containing the monthly prices. 

```{r}
aapl_prices %>%
    tq_transform(x_fun = OHLCV, transform_fun = to.period, period = "months")
```


#### tq_mutate

The brother of `tq_transform()` is `tq_mutate()`. While `tq_transform()` produces a new, transformed data set, `tq_mutate()` modifies the existing data set. Let's use `tq_mutate()` to add some Bollinger Bands and MACD using the closing prices (Cl in OHLC notation).

```{r}
aapl_prices %>%
    tq_mutate(x_fun = Cl, mutate_fun = MACD) %>%
    tq_mutate(x_fun = HLC, mutate_fun = BBands)
```

## Further Information

See the [`tidyqaunt` vignette](vignettes/tidyquant.md) for further details on the package.

