---
title: "R Quantitative Analysis Package Integrations in tidyquant"
author: "Matt Dancho"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{R Quantitative Analysis Package Integrations in tidyquant}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      fig.width = 6)
# devtools::load_all() # Travis CI fails on load_all()
```

> Functions that leverage the quantitative analysis functionality of `xts`, `zoo`, `quantmod` and `TTR`

# Overview

There's a wide range of useful quantitative analysis functions that work with time-series objects. The problem is that many of these _wonderful_ functions don't work data frames. That is until now! The `tidyquant` package integrates the most useful functions from the `xts`, `zoo`, `quantmod` and `TTR` packages. This vignette focuses on the following _core functions_ to demonstrate how the integratation works with the quantitative finance packages:

* Transmute, `tq_transmute()`: Returns a new tidy data frame typically in a different periodicity than the input.
* Mutate, `tq_mutate()`: Adds columns to the existing tidy data frame.


# Prerequisites

Load the `tidyquant` package to get started.

```{r}
# Loads tidyquant, tidyverse, lubridate, xts, quantmod, TTR 
library(tidyquant)
```

# Function Compatibility

`tq_transmute_fun_options()` returns a list the __compatible mutate functions__ by each package. We'll discuss these options by package briefly.

```{r}
tq_transmute_fun_options() %>% str()
```

## zoo Functionality

```{r}
# Get zoo functions that work with tq_transmute and tq_mutate
tq_transmute_fun_options()$zoo
```

The `zoo` functions that are compatible are listed above. Generally speaking, these are the:

* Roll Apply Functions:
    * A generic function for applying a function to rolling margins.
    * Form: `rollapply(data, width, FUN, ..., by = 1, by.column = TRUE, fill = if (na.pad) NA, na.pad = FALSE, partial = FALSE, align = c("center", "left", "right"), coredata = TRUE)`.
    * Options include `rollmax`, `rollmean`, `rollmedian`, `rollsum`, etc.


## xts Functionality


```{r}
# Get xts functions that work with tq_transmute and tq_mutate
tq_transmute_fun_options()$xts
```

The `xts` functions that are compatible are listed above. Generally speaking, these are the:

* Period Apply Functions:
    * Apply a function to a time segment (e.g. `max`, `min`, `mean`, etc).
    * Form: `apply.daily(x, FUN, ...)`.
    * Options include apply.daily, weekly, monthly, quarterly, yearly.

* To-Period Functions:
    * Convert a time series to time series of lower periodicity (e.g. convert daily to monthly periodicity).
    * Form: `to.period(x, period = 'months', k = 1, indexAt, name = NULL, OHLC = TRUE, ...)`.
    * Options include to.minutes, hourly, daily, weekly, monthly, quarterly, yearly.
    * __Note 1 (Important)__: The return structure is different for `to.period` and the `to.monthly` (`to.weekly`, `to.quarterly`, etc) forms. `to.period` returns a date, while `to.months` returns a character MON YYYY. Best to use `to.period` if you want to work with time-series via `lubridate`.  
     
## quantmod Functionality

```{r}
# Get quantmod functions that work with tq_transmute and tq_mutate
tq_transmute_fun_options()$quantmod
```

The `quantmod` functions that are compatible are listed above. Generally speaking, these are the:

* Percentage Change (Delt) and Lag Functions
    * Delt: `Delt(x1, x2 = NULL, k = 0, type = c("arithmetic", "log"))`
        * Variations of Delt: ClCl, HiCl, LoCl, LoHi, OpCl, OpHi, OpLo, OpOp 
        * Form: `OpCl(OHLC)`
    * Lag: `Lag(x, k = 1)` / Next: `Next(x, k = 1)` (Can also use `dplyr::lag` and `dplyr::lead`)
    

* Period Return Functions: 
    * Get the arithmetic or logarithmic returns for various periodicity, which include daily, weekly, monthly, quarterly, and yearly.
    * Form: `periodReturn(x, period = 'monthly', subset = NULL, type = 'arithmetic', leading = TRUE, ...)`

* Series Functions: 
    * Return values that describe the series. Options include describing the increases/decreases, acceleration/deceleration, and hi/low.
    * Forms: `seriesHi(x)`, `seriesIncr(x, thresh = 0, diff. = 1L)`, `seriesAccel(x)`

## TTR Functionality

```{r}
# Get TTR functions that work with tq_transmute and tq_mutate
tq_transmute_fun_options()$TTR
```


Here' a brief description of the most popular functions from `TTR`:

* Welles Wilder's Directional Movement Index: 
    *  `ADX(HLC, n = 14, maType, ...)`
* Bollinger Bands: 
    *  `BBands(HLC, n = 20, maType, sd = 2, ...)`: Bollinger Bands
* Rate of Change / Momentum: 
    * `ROC(x, n = 1, type = c("continuous", "discrete"), na.pad = TRUE)`: Rate of Change
    * `momentum(x, n = 1, na.pad = TRUE)`: Momentum
* Moving Averages (maType):
    * `SMA(x, n = 10, ...)`: Simple Moving Average
    * `EMA(x, n = 10, wilder = FALSE, ratio = NULL, ...)`: Exponential Moving Average
    * `DEMA(x, n = 10, v = 1, wilder = FALSE, ratio = NULL)`: Double Exponential Moving Average
    * `WMA(x, n = 10, wts = 1:n, ...)`: Weighted Moving Average
    * `EVWMA(price, volume, n = 10, ...)`: Elastic, Volume-Weighted Moving Average
    * `ZLEMA(x, n = 10, ratio = NULL, ...)`: Zero Lag Exponential Moving Average
    * `VWAP(price, volume, n = 10, ...)`: Volume-Weighted Moving Average Price
    * `VMA(x, w, ratio = 1, ...)`: Variable-Length Moving Average
    * `HMA(x, n = 20, ...)`: Hull Moving Average
    * `ALMA(x, n = 9, offset = 0.85, sigma = 6, ...)`: Arnaud Legoux Moving Average
* MACD Oscillator: 
    *  `MACD(x, nFast = 12, nSlow = 26, nSig = 9, maType, percent = TRUE, ...)`
* Relative Strength Index: 
    *  `RSI(price, n = 14, maType, ...)`
* runFun: 
    * `runSum(x, n = 10, cumulative = FALSE)`: returns sums over a n-period moving window.
    * `runMin(x, n = 10, cumulative = FALSE)`: returns minimums over a n-period moving window.
    * `runMax(x, n = 10, cumulative = FALSE)`: returns maximums over a n-period moving window.
    * `runMean(x, n = 10, cumulative = FALSE)`: returns means over a n-period moving window.
    * `runMedian(x, n = 10, non.unique = "mean", cumulative = FALSE)`: returns medians over a n-period moving window.
    * `runCov(x, y, n = 10, use = "all.obs", sample = TRUE, cumulative = FALSE)`: returns covariances over a n-period moving window.
    * `runCor(x, y, n = 10, use = "all.obs", sample = TRUE, cumulative = FALSE)`: returns correlations over a n-period moving window.
    * `runVar(x, y = NULL, n = 10, sample = TRUE, cumulative = FALSE)`: returns variances over a n-period moving window.
    * `runSD(x, n = 10, sample = TRUE, cumulative = FALSE)`: returns standard deviations over a n-period moving window.
    * `runMAD(x, n = 10, center = NULL, stat = "median", constant = 1.4826, non.unique = "mean", cumulative = FALSE)`: returns median/mean absolute deviations over a n-period moving window.
    * `wilderSum(x, n = 10)`: retuns a Welles Wilder style weighted sum over a n-period moving window.
* Stochastic Oscillator / Stochastic Momentum Index:
    * `stoch(HLC, nFastK = 14, nFastD = 3, nSlowD = 3, maType, bounded = TRUE, smooth = 1, ...)`: Stochastic Oscillator
    * `SMI(HLC, n = 13, nFast = 2, nSlow = 25, nSig = 9, maType, bounded = TRUE, ...)`: Stochastic Momentum Index


# Quantitative Power In Action

We'll go through some examples, but first let's get some data. The default for `tq_get()` is `get = "stock.prices"`, so all we need is to give `x` a stock symbol.

```{r}
AAPL <- tq_get("AAPL")
```

## Example 1: Changing the Periodicity from Daily to Weekly

The `xts::to.period` function is used for periodicity aggregation (converting from a lower level periodicity to a higher level such as minutes to hours or months to years). Because we are seeking a return structure that is on a different time scale than the input (daily versus weekly), we need to use a transmute function. We select `tq_transmute` and pass the open, high, low, close and volume using OHLC format via `ohlc_fun = OHLCV`. Looking at the documentation for `to.period`, we see that it accepts a `period` argument that we can set to `"weeks"`. The result is the OHLCV data returned with the dates changed to one day per week.

```{r}
AAPL %>%
    tq_transmute(ohlc_fun = OHLCV, mutate_fun = to.period, period = "weeks")
```

As a side note, we could also do columns individually using the `tq_transmute_xy` form. This is really useful for columns that don't fit the OHLC notation such as "adjusted" prices. Here's how to implement to get the adjusted prices on a weekly periodicity.

```{r}
AAPL %>%
    tq_transmute_xy(x = adjusted, mutate_fun = to.period, period = "weeks")
```



## Example 2: Getting the Max Close Price for Each Quarter.

The `xts::apply.quarterly()` function that is part of the period apply group can be used to apply functions by quarterly time segments. Because we are seeking a return structure that is on a different time scale than the input (quarterly versus daily), we need to use a transmute function. We select `tq_transmute` and pass the close price using OHLC format via `ohlc_fun = Cl`, and we send this subset of the data to the `apply.quarterly` function via the `mutate_fun` argument. Looking at the documentation for `apply.quarterly`, we see that we can pass a function to the argument, `FUN`. We want the maximum values, so we set `FUN = max`. The result is the quarters returned as a date and the maximum closing price during the quarter returned as a double. 

```{r}
AAPL %>%
    tq_transmute(ohlc_fun = Cl, mutate_fun = apply.quarterly, FUN = max)
```

Note that as an alternative you could use the xy form, replacing `ohlc_fun = Cl` with `x = close`.

## Example 3: Getting Returns 

The `quantmod::periodReturn()` function generates returns by periodicity. We'll go through a couple usage cases.  

### Example 3A: Getting and Charting Annual Returns

We want to use the adjusted closing prices column (adjusted for stock splits, which can make it appear that a stock is performing poorly if a split is included). We set `ohlc_fun = Ad`. We research the `periodReturn` function, and we found that it accepts `type = "arithmetic"` and `period = "yearly"`, which returns the annual returns. 

```{r}
AAPL_annual_returns <- AAPL %>%
    tq_transmute(ohlc_fun = Ad, mutate_fun = periodReturn, 
                 period = "yearly", type = "arithmetic")
AAPL_annual_returns
```

Charting annual returns is just a quick use of the `ggplot2` package. We filter out the 2017 date since this year is not currently complete. 

```{r}
AAPL_annual_returns %>%
    filter(year(date) < 2017) %>%
    ggplot(aes(x = date, y = yearly.returns, col = yearly.returns)) +
    geom_hline(yintercept = 0, color = "darkblue") +
    geom_segment(aes(xend = date, yend = 0)) +
    geom_point(size = 5) +
    scale_color_gradient(low = "red", high = "darkblue") +
    scale_y_continuous(limits = c(-2, 2), labels = scales::percent) +
    labs(title = "AAPL: Annual Returns",
         subtitle = "Get annual returns quickly with tq_transmute!",
         y = "Annual Returns", x = "")
```



### Example 3B: Getting Daily Log Returns

Daily log returns follows a similar approach. Normally I go with a transmute function, `tq_transmute`, because the `periodReturn` function accepts different periodicity options, and anything other than daily will blow up a mutation. But, in our situation the period returns periodicity is the same as the stock prices periodicity (both daily), so we can use either. We want to use the adjusted closing prices column (adjusted for stock splits, which can make it appear that a stock is performing poorly if a split is included), so we set `ohlc_fun = Ad`. We researched the `periodReturn` function, and we found that it accepts `type = "log"` and `period = "daily"`, which returns the daily log returns. 


```{r}
AAPL %>%
    tq_transmute(ohlc_fun = Ad, mutate_fun = periodReturn, 
                 type = "log", period = "daily")
```

## Example 4: Adding MACD and Bollinger Bands to a OHLC data set

In reviewing the available options in the `TTR` package, we see that `MACD` and `BBands` functions will get us where we need to be. In researching the documentation, the return is in the same periodicity as the input and the functions work with OHLC functions, so we can use `tq_mutate()`. MACD requires a price, so we select close using `Cl`, BBands requires high, low, and close, prices so we use `HLC`. We can chain the inputs together using the pipe (`%>%`) since mutate just adds columns. The result is a tibble containing the MACD and Bollinger Band results. 

```{r}
AAPL %>%
    tq_mutate(Cl, MACD) %>%
    tq_mutate(HLC, BBands)
```

Note that for the MACD, we could have used `tq_mutate_xy()`, setting `x = close`. However, for the BBands, we are forced to use `tq_mutate()` because of the HLC input.

## Example 5: Getting the Percentage Difference Between Open and Close from Zero to Five Periods 

We can't use the `OpCl` function for this task since it only returns the percentage difference for a period lag of zero. We keep digging and we find the base `Delt` function from quantmod. In researching the function, we see that `Delt` takes one or two inputs, `k` a series of lags, and the type of difference, either arithmetic or log. We will set `x = open` and `y = close` and `k = 0:5` to get zero through five periods. The default `type = "arithmetic"` is acceptable, so there is no need to specify. The result is the percentage difference between the open and close prices for periods zero to five.

```{r}
AAPL %>%
    tq_mutate_xy(x = open, y = close, mutate_fun = Delt, k = 0:5) %>%
    select(-c(high, low, volume, adjusted))
```

For comparison we'll inspect the output from the `OpCl()` function using `tq_mutate()`. We send OHLC prices to the OpCl function. As expected the column returned is the same as `Delt.0.arithmetic` from above. Note that we can use the `col_rename` argument to quickly rename the returned column to be the same as the `Delt` function return. 

```{r}
AAPL %>%
    tq_mutate(OHLC, OpCl, col_rename = "Delt.0.arithmetic") %>%
    select(-c(high, low, volume, adjusted))
```

## Example 6: Get the 5, 10, 15-Day Rolling Minimum and Maximum Values of the Adjusted Prices

Rolling functions come from the `zoo` package. In reviewing the available options, we see that `rollmax` is present but no there is no `rollmin`. However, the generic `rollapply` function will work with any function, so let's use that. In reviewing the documentation, the `rollapply()` function takes three primary arguments: `data`, `width` and `FUN`. The `data` is passed using the `ohlc_fun` argument, so we don't need to worry about this. The `width` is the number of periods to apply the function, and for our situation this is 5, 10, and 15. The `FUN` is the function to apply, and for our situation this is `min` and `max`. We want the result to add columns to our data set, so we use the `tq_mutate()` function, which allows us to pipe (`%>%`) each mutation. Putting it all together:

```{r}
AAPL %>%
    tq_mutate(Ad, rollapply, width = 5, FUN = min, col_rename = "roll.min.5") %>%
    tq_mutate(Ad, rollapply, width = 10, FUN = min, col_rename = "roll.min.10") %>%
    tq_mutate(Ad, rollapply, width = 15, FUN = min, col_rename = "roll.min.15") %>%
    tq_mutate(Ad, rollapply, width = 5, FUN = max, col_rename = "roll.max.5") %>%
    tq_mutate(Ad, rollapply, width = 10, FUN = max, col_rename = "roll.max.10") %>%
    tq_mutate(Ad, rollapply, width = 15, FUN = max, col_rename = "roll.max.15")
```

Note that the new column names were added using the `col_rename` argument. If this was left off, the column names are auto-generated from the `mutate_fun` name with a sequential suffix added, which is likely not descriptive enough for your needs.  

