---
title: "Scaling Your Analysis with tidyquant"
author: "Matt Dancho"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Scaling Your Analysis with tidyquant}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      fig.width = 6)
# devtools::load_all() # Travis CI fails on load_all()
```

> Designed to be used and scaled with the `tidyverse`

# Overview

The greatest benefit to `tidyquant` is the ability to easily scale your financial analysis. Scaling is the process of creating an analysis for one security and then extending it to multiple groups. This is idea of scaling is incredibly useful to financial analysts because typically one wants to compare many securities to make informed decisions. Fortunately, the `tidyquant` package integrates with the `tidyverse` making scaling super simple!

All `tidyquant` functions return data in the `tibble` (tidy data frame) format, which allows for interaction within the `tidyverse`. This means we can:

* Seamlessly scale data retrieval and transformations/mutations
* Use the pipe (`%>%`) for chaining operations
* Use `dplyr` and `tidyr`: `select`, `filter`, `group_by`, `nest`/`unnest`, `spread`/`gather`, etc
* Use `purrr`: mapping functions with `map`

We'll go through some useful scaling techniques for getting and manipulating groups of data.

# Prerequisites

Load the `tidyquant` package to get started.

```{r}
# Loads tidyquant, tidyverse, lubridate, xts, quantmod, TTR 
library(tidyquant)  
```

 

# Getting Financial Data for Multiple Stocks

A very basic example is retrieving the stock prices for multiple stocks. There are three primary ways to do this:


## Method 1: Map a character vector with multiple stock symbols

```{r}
c("AAPL", "GOOG", "FB") %>%
    tq_get(get = "stock.prices", from = "2016-01-01", to = "2017-01-01")
```

The output is a single level tibble with all or the stock prices in one tibble. The auto-generated column name is "symbol", which can be pre-emptively renamed by giving the vector a name (e.g. `stocks <- c("AAPL", "GOOG", "FB")`) and then piping to `tq_get`.


## Method 2: Map a tibble with stocks in first column

First, get a stock list in data frame format either by making the tibble or retrieving from `tq_index` / `tq_exchange`. The stock symbols must be in the first column.

### Method 2A: Make a tibble

```{r}
stock_list <- tibble(stocks = c("AAPL", "JPM", "CVX"),
                     industry = c("Technology", "Financial", "Energy"))
stock_list
```

Second, send the stock list to `tq_get`. Notice how the symbol and industry columns are automatically expanded the length of the stock prices.


```{r}
stock_list %>%
    tq_get(get = "stock.prices", from = "2016-01-01", to = "2017-01-01")
```

### Method 2B: Use index or exchange

Get an index...

```{r}
tq_index("DOWJONES")
```

...or, get an exchange.

```{r}
tq_exchange("NYSE")
```


Send the index or exchange to `tq_get`. _Important Note: This can take several minutes depending on the size of the index or exchange, which is why only the first three stocks are evaluated in the vignette._

```{r}
tq_index("DOWJONES") %>%
    slice(1:3) %>%
    tq_get(get = "stock.prices")
```

You can use any applicable "getter" to get data for __every stock in an index or an exchange__! This includes: "stock.prices", "key.ratios", "key.stats", "financials", and more.


## Method 3: Use purrr to map a function

We can pipe a tibble of stock symbols to a mutation that maps the `tq_get(get = "stock.prices")` function. The result is all of the stock prices in nested format.

```{r}
tibble(symbol = c("AAPL", "GOOG", "AMZN", "FB")) %>%
    mutate(stock.prices = map(.x = symbol, ~ tq_get(.x, get = "stock.prices")))
```


## Compound Getters

In financial analysis, it's very common to need data from various sources to combine in an analysis. For this reason multiple `get` options ("compound getters") can be used to return a "compound get". A quick example:

```{r}
c("AAPL", "GOOG") %>%
    tq_get(get = c("stock.prices", "financials"))
```

This returns the stock prices and financials for each stock as one nested data frame! Any of the `get` options that accept stock symbols can be used in this manner: `"stock.prices"`, `"financials"`, `"key.ratios"`, `"key.stats"`, `"dividends"`, and `"splits"`. 

This capability becomes incredibly useful when combined with `purrr` function mapping, which is discussed in [Manipulating Financial Data with purrr](#purrr-function-mapping).


# Manipulating Financial Data using Groups

Once you get the data, you typically want to do something with it. You can easily do this at scale. Let's get the yearly returns for multiple stocks using `tq_transform`. First, get the prices. Second, use `group_by` to group by stock symbol. Third, apply the transformation. We can do this in one easy workflow:

```{r}
c("AAPL", "GOOG", "FB") %>%
    tq_get(get = "stock.prices", from = "2012-01-01", to = "2017-01-01") %>%
    group_by(symbol) %>%
    tq_transform(Ad, transform_fun = periodReturn, period = "yearly", 
                 col_rename = "yearly.returns") %>%
    ggplot(aes(x = year(date), y = yearly.returns, fill = symbol)) +
    geom_bar(position = "dodge", stat = "identity") +
    scale_y_continuous(labels = scales::percent) +
    scale_x_continuous(breaks = seq(2008, 2017, by = 1)) +
    labs(title = "AAPL, GOOG, FB: Annual Returns", 
         subtitle = "Transforming using quantmod functions is easy!", 
         x = "") +
    theme(legend.position = "bottom")
```

<a class="anchor" id="purrr-function-mapping"></a>

# Manipulating Financial Data using purrr

Eventually you will want to begin modeling at scale! One of the __best__ features of the `tidyverse` is the ability to map functions to nested tibbles using `purrr`. From the Many Models chapter of "[R for Data Science](http://r4ds.had.co.nz/)", we can apply the same modeling workflow to financial analysis. Using a two step workflow:

1. Analyze a single stock
2. Scale to many stocks

Let's go through an example to illustrate. In our hypothetical situation, we want to compare the mean monthly log returns (MMLR). 

## Analyze a Single Stock

First, let's come up with a function to help us collect log returns. The function below performs three operations internally. It first gets the stock prices using `tq_get()`. Then, it transforms the stock prices to period returns using `tq_transform()`. We add the `type = "log"` and `period = "monthly"` arguments to ensure we retrieve a tibble of monthly log returns. Last, we take the mean of the monthly returns to get MMLR.

```{r}
my_stock_analysis_fun <- function(stock.symbol) {
    period.returns <- stock.symbol %>%
        tq_get(get = "stock.prices") %>%
        tq_transform(ohlc_fun = Ad, transform_fun = periodReturn, 
                     type = "log", period = "monthly")
    mean(period.returns$monthly.returns)
}
```

And, let's test it out. We now have the mean monthly log returns over the past ten years.

```{r}
my_stock_analysis_fun("AAPL")
```



## Scale to Many Stocks 

Now that we have one stock down, we can scale to many stocks. For brevity, we'll randomly sample ten stocks from the S&amp;P500 with a call to `dplyr::sample_n()`.

```{r}
set.seed(100)
stocks <- tq_index("SP500") %>%
    sample_n(10)
stocks
```

We can now apply our analysis function to the stocks using `dplyr::mutate` and `purrr::map_dbl`. The `mutate()` function adds a column to our tibble, and the `map_dbl()` function maps our `my_stock_analysis_fun` to our tibble of stocks using the `symbol` column.

```{r}
stocks <- stocks %>%
    mutate(mmlr = map_dbl(symbol, my_stock_analysis_fun)) %>%
    arrange(desc(mmlr))
stocks
```

And, we're done! We now have the MMLR for 10-years of stock data for 10 stocks. And, we can easily extend this to larger lists or stock indexes. For example, the entire S&amp;P500 could be analyzed removing the `sample_n()` following the call to `tq_index("SP500")`.

# Error Handling when Scaling 

Eventually you will run into a stock index, stock symbol, FRED data code, etc that cannot be retrieved. Possible reasons are: 

* An index becomes out of date
* A company goes private
* A stock ticker symbol changes
* Yahoo / FRED just doesn't like your stock symbol / FRED code

This becomes painful when scaling if the functions return errors. So, the `tq_get()` function is designed to handle errors _gracefully_. What this means is an `NA` value is returned when an error is generated along with a _gentle error warning_. 

```{r}
tq_get("XYZ", "stock.prices")
```

## Pros and Cons to Built-In Error-Handling

There are pros and cons to this approach that you may not agree with, but I believe helps in the long run. Just be aware of what happens:

* __Pros__: Long running scripts are not interrupted because of one error

* __Cons__: Errors can be inadvertently handled or flow downstream if the users does not read the warnings


## Bad Apples Fail Gracefully, tq_get

Let's see an example when using `tq_get()` to get the stock prices for a long list of stocks with one `BAD APPLE`. The argument `complete_cases` comes in handy. The default is `TRUE`, which removes "bad apples" so future analysis have complete cases to compute on. Note that a gentle warning stating that an error occurred and was dealt with by removing the rows from the results.

```{r, warning = TRUE}
c("AAPL", "GOOG", "BAD APPLE") %>%
    tq_get(get = "stock.prices", complete_cases = TRUE)
```

Now switching `complete_cases = FALSE` will retain any errors as `NA` values in a nested data frame. Notice that the error message and output change. The error message now states that the `NA` values exist in the output and the return is a "nested" data structure. 

```{r, warning = TRUE}
c("AAPL", "GOOG", "BAD APPLE") %>%
    tq_get(get = "stock.prices", complete_cases = FALSE)
```

In both cases, the prudent user will review the warnings to determine what happened and whether or not this is acceptable. In the `complete_cases = FALSE` example, if the user attempts to perform downstream computations at scale, the computations will likely fail grinding the analysis to a hault. But, the advantage is that the user will more easily be able to filter to the problem childs to determine what happened and decide whether this is acceptable or not.
